Features for detecting threats of violence in YouTube comments

### 2016-03-30 submission.tex
Changed formatting of TeX code

\documentclass[11pt,letterpaper]{article}
\usepackage{naaclhlt2016}
\usepackage{times}
\usepackage{latexsym}
\usepackage{relsize}

@@

%\naaclfinalcopy % Uncomment this line for the final submission
\def\naaclpaperid{***} %  Enter the naacl Paper ID here

@@

% Author information can be set in various styles:
% For several authors from the same institution:
% \author{Author 1 \and ... \and Author n \\
%         Address line \\ ... \\ Address line}
% if the names do not fit well on one line use

%         Author 1 \\ {\bf Author 2} \\ ... \\ {\bf Author n} \\
% For authors from different institutions:
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \And  ... \And
%         Author n \\ Address line \\ ... \\ Address line}
% To start a seperate ``row'' of authors use \AND, as in
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \AND
%         Author 2 \\ Address line \\ ... \\ Address line \And
%         Author 3 \\ Address line \\ ... \\ Address line}
% If the title and author information does not fit in the area allocated,
% place \setlength\titlebox{<new height>} right after
% at the top, where <new height> can be something larger than 2.25in
\author{Author 1\\
	    XYZ Company\\
	    111 Anywhere Street\\
	    Mytown, NY 10000, USA\\
	    {\tt author1@xyz.org}
	  \And
	Author 2\\
  	ABC University\\
  	900 Main Street\\
  	Ourcity, PQ, Canada A1A 1T2\\
  {\tt author2@abc.ca}}

@@

\begin{abstract}
This article investigates the effect of various types of linguistic
features (lexical, morphosyntactic and lexical semantic) for the task of
detecting threats of violence in a data set of YouTube comments. It
further examines the use of these features in combination with
different representations of linguistic context, and feature backoff. Our results show that combinations of lexical features outperform the use of more complex syntactic and semantic features for this task.
\end{abstract}

@@

Perhaps closest to the current work is the work of
\newcite{hammer2014}, which reports on an experiment on the same data set that
we will be using in our own experiments. The method uses a logistic
LASSO regression analysis on bigrams (skip-grams) of important words
to classify sentences as either \tovs{} or not.  The system makes use of a
list of words that are correlated with threats of violence. The
article does not, however, describe exactly how these important words
were selected, stating only that words were chosen that were
significantly correlated with the response (violent/non-violent
sentence). Results are reported only in terms of proportion of false
positives for the two classes, and it is not clear how the data was
split for training and evaluation. This makes it difficult to compare our results to those of \newcite{hammer2014}.

\section{The YouTube threat data set}
\label{sec:data}
The YouTube threat data set is comprised of user-written comments from
eight different YouTube videos \cite{hammer2014}. A comment
consists of a set of sentences, each of them manually annotated to be
either a \tov{} (or support for a \tov{}), or not. The data set
furthermore records the username
%% \footnote{In 2013, YouTube changed its
%%   commenting system from using unique usernames, to using "real
%%   names", like Facebook and other sites \cite{youtube2013}. Some
%%   accounts, however, did not provide real names, so they continue to
%%   only be identified by their usernames.}
  of the user that posted the
comment. The eight videos that the comments were posted to cover religious and political topics like halal slaughter, immigration,
Anders Behring Breivik, Jihad, etc. \cite{hammer2014}. 
%The videos cover slightly different topics, the videos all contain the same
%type of discussions, namely of xenophobia and racism.

The data set consists of 9,845 comments, comprised of 28,643
sentences, see Table \ref{tab:dataset}.
%% In total there are 402,673 tokens in the sentences in the data set.
There are 1,285 comments
containing threats, and 1,384 sentences containing threats, as seen in
Table \ref{tab:dataset}.  \newcite{hammer2014} reports inter-annotator
agreement on this data set to be 98~\%, as calculated on 120 of the
comments, doubly annotated for evaluation.

@@

%to-do: should we anonymize the user names?
Figure \ref{fig:comments} provides some examples of comments containing
\tovs{} taken from the data set. The first line is the annonymized username, and the subsequent lines are the sentences of the comment.
%% An empty line indicates the end of a comment.
The sentences are annotated
with a number indicating whether they contain a \tov{} (1), or not
(0).

@@

User #88
0    Need a solution?
1    Drop one good ol' nuke on that
     black toilet in Mecca.
\end{Verbatim}
\caption{Examples of comments from the data set.}
\label{fig:comments}
\end{figure}

@@

\paragraph{Pre-processing}Since threat annotation is performed on the sentence level, the data
set has been manually split into sentences as part of the annotation
process.  We performed tokenization, lemmatization, PoS-tagging and
dependency parsing using the SpaCy
toolkit\footnote{https://spacy.io/}.  SpaCy assigns both the standard
Penn Treebank PoS-tags \cite{Mar:San:Mar:93}, as well as the more
coarse-grained Universal PoS tag set of \newcite{Pet:Das:McD:12}. The
dependency parser assigns an analysis in compliance with the ClearNLP
converter \cite{Cho:Pal:12}, see Figure \ref{fig:dep} for an example
dependency graph from the data set. The data set was further enriched
with the cluster labels described in Turian et
al.~\shortcite{Tur:Rat:Ben:10}, created using the Brown clustering
algorithm \cite{Bro:deS:Mer:92} and induced from the RCV1 corpus, a
corpus containing Reuters English newswire text.
%, with approximately 63 million words and 3.3 million sentences.
We vary the number of
clusters to be either 100, 320, 1000 or 3200 clusters and use the full
cluster label.  We also make use of the WordNet resource
\cite{wordnet} to determine a word's synset, parent and grandparent
synset.

@@

\begin{figure}
\begin{center}
\begin{dependency}
%\begin{dependency}[arc edge, text only label, theme=simple]
  \begin{deptext}[column sep=.2cm]
    i \& m \& fucking \& scared \& kill \& them \& d\\
   \tag{\smaller pron} \& \tag{\smaller verb} \& \tag{\smaller verb} \& \tag{\smaller adj} \& \tag{\smaller verb} \& \tag{\smaller pron} \& \tag{\smaller x} \\
  \end{deptext}
  \depedge{2}{1}{\large nsubj}
  %% \depedge[edge unit distance=2ex]{2}{8}{\large punct}
  \depedge{2}{4}{\large acomp}
  \depedge{4}{3}{\large amod}
  \depedge{2}{5}{\large conj}
  \depedge{5}{6}{\large dobj}
  \depedge{5}{7}{\large advmod}
 \deproot{2}{root}
\end{dependency}
\caption{Dependency parse of example sentence from the data set, with assigned uPOS tags.}
\label{fig:dep}
\end{center}
\end{figure}

@@

\paragraph{Features}
Based on the enriched data set, as described above, we experiment with the following feature types:

>>>>

\documentclass[11pt,letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage{naaclhlt2016}
\usepackage{times}
\usepackage{latexsym}
\usepackage{relsize}

@@

\naaclfinalcopy % Uncomment this line for the final submission
\def\naaclpaperid{***} %  Enter the naacl Paper ID here

@@

% Author information can be set in various styles:
% For several authors from the same institution:
\author{Aksel Wester \and Lilja Øvrelid \and Erik Velldal \\
        University of Oslo \\ ... \\ Address line}
% if the names do not fit well on one line use

%         Author 1 \\ {\bf Author 2} \\ ... \\ {\bf Author n} \\
% For authors from different institutions:
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \And  ... \And
%         Author n \\ Address line \\ ... \\ Address line}
% To start a seperate ``row'' of authors use \AND, as in
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \AND
%         Author 2 \\ Address line \\ ... \\ Address line \And
%         Author 3 \\ Address line \\ ... \\ Address line}
% If the title and author information does not fit in the area allocated,
% place \setlength\titlebox{<new height>} right after
% at the top, where <new height> can be something larger than 2.25in
%% \author{Aksel Wester\\
%%   University of Oslo\\
%%   111 Anywhere Street\\
%%   Mytown, NY 10000, USA\\
%%   {\tt aksellw@ifi.uio.no}
%%   \And
%%   Lilja Øvrelid\\
%%   University of Oslo\\
%%   900 Main Street\\
%%   Ourcity, PQ, Canada A1A 1T2\\
%%   {\tt liljao@ifi.uio.no}
%%   \And
%%   Erik Velldal\\
%%   University of Oslo\\
%%   900 Main Street\\
%%   Ourcity, PQ, Canada A1A 1T2\\
%%   {\tt erikve@ifi.uio.no}}

@@

\begin{abstract}
This article investigates the effect of various types of linguistic features (lexical, morphosyntactic and lexical semantic) for the task of detecting threats of violence in a corpus of YouTube comments. It further examines the use of these features in combination with different representations of linguistic context, and feature backoff. Our results show that combinations of lexical features outperform the use of more complex syntactic and semantic features for this task.
\end{abstract}

@@

Perhaps closest to the current work is the work of \newcite{hammer2014}, which reports on an experiment on the same corpus that we will be using in our own experiments. The method uses a logistic LASSO regression analysis on bigrams (skip-grams) of important words to classify sentences as either \tovs{} or not.  The system makes use of a list of words that are correlated with threats of violence. The article does not, however, describe exactly how these important words were selected, stating only that words were chosen that were significantly correlated with the response (violent/non-violent sentence). Results are reported only in terms of proportion of false positives for the two classes, and it is not clear how the data was split for training and evaluation. This makes it difficult to compare our results to those of \newcite{hammer2014}.

\section{The YouTube threat corpus}
\label{sec:data}
The YouTube threat corpus is comprised of user-written comments from eight different YouTube videos \cite{hammer2014}. A comment consists of a set of sentences, each of them manually annotated to be either a \tov{} (or support for a \tov{}), or not. The corpus furthermore records the username
%% \footnote{In 2013, YouTube changed its
%%   commenting system from using unique usernames, to using "real
%%   names", like Facebook and other sites \cite{youtube2013}. Some
%%   accounts, however, did not provide real names, so they continue to
%%   only be identified by their usernames.}
  of the user that posted the comment. The eight videos that the comments were posted to cover religious and political topics like halal slaughter, immigration, Anders Behring Breivik, Jihad, etc. \cite{hammer2014}. 
%The videos cover slightly different topics, the videos all contain the same
%type of discussions, namely of xenophobia and racism.

The corpus consists of 9,845 comments, comprised of 28,643 sentences, see Table \ref{tab:dataset}.
%% In total there are 402,673 tokens in the sentences in the data set.
There are 1,285 comments containing threats, and 1,384 sentences containing threats, as seen in Table \ref{tab:dataset}.  \newcite{hammer2014} reports inter-annotator agreement on this data set to be 98~\%, as calculated on 120 of the comments, doubly annotated for evaluation.

@@

%to-do: should we anonymize the user names?
Figure \ref{fig:comments} provides some examples of comments containing \tovs{} taken from the corpus. The first line is the annonymized username, and the subsequent lines are the sentences of the comment. %% An empty line indicates the end of a comment.
The sentences are annotated with a number indicating whether they contain a \tov{} (1), or not (0).

@@

User #88
0    Need a solution?
1    Drop one good ol' nuke on that
     black toilet in Mecca.
\end{Verbatim}
\caption{Examples of comments from the YouTube threat corpus.}
\label{fig:comments}
\end{figure}

@@

\paragraph{Pre-processing}Since threat annotation is performed on the sentence level, the corpus has been manually split into sentences as part of the annotation process.  We performed tokenization, lemmatization, PoS-tagging and dependency parsing using the SpaCy toolkit\footnote{https://spacy.io/}.  SpaCy assigns both the standard Penn Treebank PoS-tags \cite{Mar:San:Mar:93}, as well as the more coarse-grained Universal PoS tag set of \newcite{Pet:Das:McD:12}. The dependency parser assigns an analysis in compliance with the ClearNLP converter \cite{Cho:Pal:12}, see Figure \ref{fig:dep} for an example dependency graph from the corpus. The corpus was further enriched with the cluster labels described in Turian et al.~\shortcite{Tur:Rat:Ben:10}, created using the Brown clustering algorithm \cite{Bro:deS:Mer:92} and induced from the RCV1 corpus, a corpus containing Reuters English newswire text.
%, with approximately 63 million words and 3.3 million sentences.
We vary the number of clusters to be either 100, 320, 1000 or 3200 clusters and use the full cluster label.  We also make use of the WordNet resource \cite{wordnet} to determine a word's synset, parent and grandparent synset.

@@

\begin{figure}
\begin{center}
\begin{dependency}
%\begin{dependency}[arc edge, text only label, theme=simple]
  \begin{deptext}[column sep=.2cm]
    i \& m \& fucking \& scared \& kill \& them \& d\\
   \tag{\smaller pron} \& \tag{\smaller verb} \& \tag{\smaller verb} \& \tag{\smaller adj} \& \tag{\smaller verb} \& \tag{\smaller pron} \& \tag{\smaller x} \\
  \end{deptext}
  \depedge{2}{1}{\large nsubj}
  %% \depedge[edge unit distance=2ex]{2}{8}{\large punct}
  \depedge{2}{4}{\large acomp}
  \depedge{4}{3}{\large amod}
  \depedge{2}{5}{\large conj}
  \depedge{5}{6}{\large dobj}
  \depedge{5}{7}{\large advmod}
 \deproot{2}{root}
\end{dependency}
\caption{Dependency parse of example sentence from the corpus, with assigned uPOS tags.}
\label{fig:dep}
\end{center}
\end{figure}

@@

\paragraph{Features}
Based on the enriched corpus, as described above, we experiment with the following feature types:
